---
title: "Dataframes In Class lab"
author: "Richard Sauve"
date: "DATA 2401"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Working with built-in data sets

Load R's "USPersonalExpenditure" dataset using the `data()` function

This will produce a data frame called `USPersonalExpenditure`
```{r}
data("USPersonalExpenditure")
```

The variable `USPersonalExpenditure` is now accessible to you. Unfortunately, it's not a data frame (it's actually what is called a matrix) 

Test this using the `is.data.frame()` function:

```{r}
is.data.frame(USPersonalExpenditure)
```

Luckily, you can pass the USPersonalExpenditure variable as an argument to the `data.frame()` function to convert it a data frame.

Do this, storing the result in a new variable

```{r}
USPE <- data.frame(USPersonalExpenditure)
```


What are the column names of your dataframe?
```{r}
colnames(USPE)
```


Consider: why are they so strange? (Think about whether you could use a number like 1940 with dollar notation!)

What are the row names of your dataframe?
```{r}
rownames(USPE)
```


Add a column "category" to your data frame that contains the rownames

```{r}
USPE$category <- rownames(USPE)
```


How much money was spent on personal care in 1940?
```{r}
USPE$X1940[USPE$category == "Personal Care"]
```


How much money was spent on Food and Tobacco in 1960?
```{r}
USPE$X1960[USPE$category == "Food and Tobacco"]

```


What was the highest expenditure category in 1960?
```{r}
USPE$category[which.max(USPE$X1960)]

```


Define a function `lowest_category` that takes in a year as a parameter, and returns the lowest spending category of that year
```{r}
colnames(USPE) <- c("1940","1945","1950","1955","1960","category")
print(USPE)
```

```{r}
lowest_category <- function(year){
  for (i in (1:5)){
    if (colnames(USPE[i]) == year){
      return(USPE$category[min(USPE[i]) == USPE[i]])
    }
  }
}

lowest_category(1940)
```


Using your function, determine the lowest spending category of each year
Hint: use the `sapply()` function to apply your function to a vector of years

```{r}
years <- c(1940,1945,1950,1955,1960)
lowest_years <- sapply(years, lowest_category)
print(lowest_years)
```
