---
title: "Homework 2 - Intro to R"
author: "Richard Sauve"
date: "DATA 2401"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create a variable `hometown` that stores the city in which you were born

```{r Q1}
hometown <- "Calgary"
```


2. Assign your name to the variable `my_name`

```{r Q2}
my_name <- "Richard_Sauve"
```

3. Assign your height (in inches) to a variable `my_height`

```{r Q3}
my_height <- 72
```

4. Create a variable `puppies` equal to the number of puppies you'd like to have

```{r Q4}
puppies <- 7
```

5. Create a variable `puppy_price`, which is how much you think a puppy costs

```{r Q5}
puppy_cost <- 168
```

6. Create a variable `total_cost` that has the total cost of all of your puppies (use your variables created in the previous questions)

```{r Q6}
total_cost <- puppies*puppy_cost
```

7. Create a boolean variable `too_expensive`, set to TRUE if the cost is greater than $1,000 (use your variables created in the previous questions)

```{r Q7}
too_expensive <- total_cost > 1000
```

8. Create a variable `max_puppies`, which is the number of puppies you can afford for $1,000

```{r Q8}
max_puppies <- 1000/puppy_cost
```

9.  We can't have partial puppies. Round `max_puppies` down to a whole number (you might use a function called `floor`, type `?floor` in your R console for more information), save that as a new variable `actual_max_puppies`

```{r Q9}
acutal_max_puppies <- floor(max_puppies)
```

