---
title: "In Class Lab - Lists"
author: "Richard Sauve"
date: "DATA 2401"
output:
  html_document: default
  pdf_document: default
---

**Note: Your exam will be in a similar format to this (RMarkdown File with Questions and Code chunks for your answers), so make SURE you have the process for knitting to html or pdf working.** 

This lab is designed to be if we finish early on List day OR as part of the review on Monday. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# uncomment these, run them ONCE, then recomment them
#install.packages(tidyverse)
#install.packages(ggplot2)


library(tidyverse)
library(ggplot2)

```

## Indexing 

1. The below code chunk saves some built-in data about diamonds to vectors in your environment. (Run `? diamonds` in your console for more information, you may need to install and load the `tidyverse` and `ggplot2` libraries using the code in the chunk above). Using indexing and the tools you've learned, find the following:

   - What's the mean price of diamonds? The mean carat value? 
   - How many diamonds are more than 4 carats?
   - What's the largest diamond?
   - What's the most expensive diamond? 
   - Assuming that these are paired information (`carats[1]` and `price[1]` are information about the *same diamond*), how many carats is the most expensive diamond? 

```{r}
carats <- diamonds$carat
price <- diamonds$price


```

```{r}
# What's the mean price of diamonds? The mean carat value? 
print(mean(price))
print(mean(carats))

# How many diamonds are more than 4 carats?
carats_4 <- carats > 4
print(length(carats_4))

# What's the largest diamond?
print(max(carats))

# What's the most expensive diamond? 
print(max(price))

# Assuming that these are paired information 
# (`carats[1]` and `price[1]` are information about the same diamond), 
# how many carats is the most expensive diamond? 

print(max(carats)) # isnt this the same as questions 3?
 
# find average price of diamonds under 1 carat
carats_1 <- carats < 1
print(mean(carats_1))

# what is size of the largest diamond I can get for less than 1000
diamond_1000 <- carats[price < 1000]
print(max(diamond_1000))

```

##  Functions + Indexing

1. Write a function that takes in a numeric vector, changes any negative values to 0, and returns the changed vector. 

```{r}
no_negatives <- function(vector){
   vector_pos <- c()
   if (is.vector(vector) & is.numeric(vector)){
      for (x in 1:length(vector)){
         if (vector[x] < 0){
            vector_pos <- append(vector_pos, vector[x]*-1)
         }else{
            vector_pos <- append(vector_pos, vector[x])
         }
      }
      return(vector_pos)
   }else{
      print("You did not input a numeric vector")
   }
}

my_vector <- c((5:1)*-1, 0 , 1:5)

no_negatives(my_vector)

```






##  Lists

Back to the diamond data. 

1. Make a boolean vector called `too_much` that is T/F to indicate if each element of `price` is larger than the median value of `price`.

```{r}
too_much <- price > mean(price)

```


1. Put this vector, along with `carats` and `price` in a list called `diamond_info`. What are the different ways you can access the information inside this list? 

```{r}
diamond_info <- list(carats, price)
```

```{r, eval = T}
# I set eval = F here in the chunk so these don't run and take 200+ pages... 
print(diamond_info$carats)
```

```{r, eval = T}
# let's look at what happens here
# back to eval = T, which is the default
print(diamond_info$carats)

```


1. Use `lapply` to find the mean of each variable. 
Sidenote challenge question: why is the mean value of `too_much` what it is? 

```{r}
lapply(too_much, mean)[[1]] # It was annoying having it all print out
# too_much is a list of Boolean values with True/False corresponding to 1/0
print(mean(too_much))

```

1. How can I remove the element `too_much` from the list?

```{r}
# for example if I wanted to remove the second element I put a negative in fron tof the number in the idex slow
# too_much <- too_much[-2]
```



```{r}
# price of the diamonds that are larger than 4 carats
diamond_info$my_price[ diamond_info$my_carats > 4]

# price of the diamonds that are larger than 4 carats or smaller than .3 carats
index_vec <-  diamond_info$my_carats < .3 | diamond_info$my_carats > 4
diamond_info$my_price[index_vec]
```

