---
title: "Homework 3"
author: "Richard Sauve"
date: "Due 10/18/21"
output: html_document
---

Please submit a zip file containing the *knitted* html of your answers, titled HW3_lastname.html 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Working with a created data frame

1. Create a vector of 100 employees ("Employee 1", "Employee 2", ... "Employee 100")
Hint: use the `paste()` function and vector recycling to add a number to the word
"Employee"

```{r}
employees <- c(1:100)
employees <- paste("Employee", employees)
```


2. Create a vector of 100 random salaries for the year 2017
Use the `runif()` function to pick random numbers between 40000 and 50000

```{r}
salaries_2017 <- runif(100, 40000, 50000)
```


3. Create a vector of 100 annual salary adjustments between -5000 and 10000.
(A negative number represents a salary decrease)
Again use the `runif()` function to pick 100 random numbers in that range.

```{r}
salary_adjustments <- round(runif(100, -5000, 10000), 2)

```


4. Create a data frame `salaries` by combining the 3 vectors you just made
Remember to set `stringsAsFactors=FALSE`!

```{r}
salaries <- data.frame(employees, salaries_2017, salary_adjustments, stringsAsFactors = FALSE)

```

5. Add a column to the `salaries` data frame that represents each person's
salary in 2018 (e.g., with the salary adjustment added in).

```{r}
salaries$salaries_2018 <- salaries$salaries_2017 + salaries$salary_adjustments
```


6. Add a column to the `salaries` data frame that has a value of `TRUE` if the 
person got a raise (their salary went up)

```{r}
salaries$Got_Raise <- salaries$salaries_2018 > salaries$salaries_2017
```


Retrieve values from your data frame to answer the following questions

Note that you should get the value as specific as possible (e.g., a single
cell rather than the whole row!). 

1. What was the 2018 salary of Employee 57

```{r}
salaries$salaries_2018[57]
```



2. How many employees got a raise?

```{r}
salaries %>%
  select(Got_Raise) %>%
  filter(Got_Raise == TRUE) %>%
  count()
```


3. What was the dollar value of the highest raise?

```{r}
salaries %>%
  select(salary_adjustments) %>%
  max()
```


4. What was the "name" (number) of the employee who received the highest raise?

```{r}
salaries$employees[salaries$salary_adjustments == max(salary_adjustments)]
# wow employee 57 got the highest raise in my first iteration
```



5. What was the largest decrease in salaries between the two years?

```{r}
min(salaries$salary_adjustments)
```



6. What was the name of the employee who received largest decrease in salary?

```{r}
salaries$employees[salaries$salary_adjustments == min(salary_adjustments)]
```



7. What was the average salary change?

```{r}
mean(salaries$salary_adjustments)
```



8. For people who did not get a raise, how much money did they lose on average?

```{r}
salaries %>%
  select(salary_adjustments) %>%
  filter(salary_adjustments < 0) %>%
  summarise(mean = mean(salary_adjustments)) # why cant you just pass it into mean(), didn't really understand why
```



Consider: do the above averages match what you expected them to be based on how you generated the salaries?





## Baby Name Popularity Over Time

Read in the female baby names data file found on Blackboard into a 
variable called `names`. Remember to NOT treat the strings as factors!


Create a data frame `names_2013` that contains only the rows for the year 2013
```{r}
names <- data.frame(read.csv("female_names.csv"), stringsAsFactors = FALSE)
names_2013 <- filter(names, year == 2013)
```

1. What was the most popular female name in 2013?
```{r}
names_2013 %>%
  slice_max(proportion)
```

2. Write a function `most_popular_in_year` that takes in a year as a value and 
returns the most popular name in that year
```{r}
most_popular_in_year <- function(input_year){
  names %>%
    filter(year == input_year) %>%
    slice_max(proportion)
}
```

3. Use your new function. What was the most popular female name in 1994?
```{r}
most_popular_in_year(1994)
```

4. Write a function `number_in_million` that takes in a name and a year, and 
returns statistically how many babies out of 1 million born that year have 
that name. 
Hint: get the popularity percentage, and take that percentage out of 1 million.
```{r}
number_in_million <- function(input_name, input_year){
  names %>%
    filter(name == input_name, year == input_year) %>%
    summarise(per_million = proportion*1000000)
}
```

5. How many babies out of 1 million had the name 'Laura' in 1995?
```{r}
number_in_million("Laura", 1995)
```

6. How many female babies out of 1 million had your name in the year you were born?
```{r}
number_in_million("Richard", 1998)
```

Consider: what does this all tell you about how easy it is to identify someone with just your name and birth year?
```
With only a couple more points of information, you could easily track people down or at least get the list down 
to a trial and error situation.
```
